Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-14 doc/report/msc-report.tex
doc/report/msc-report.tex: * added listings package, used for code listing. * WIP on Implementation chapter.

\documentclass[a4paper]{report}
\usepackage{graphicx}

@@

\section{Configuration}

The Zohmg needs configuration of the user's data sets and the environment for
Hadoop and Dumbo.

The data sets are configured using YAML\footnote{YAML Ain't a Markup Language},
a human-readable data serializaton format. Any human-readable configure file
method could have been chosen, e.g. INI files\footnote{A simple configuration
mechanims where one would have lines of \texttt{variable = value}.}. YAML was
choosen because it is is very simple for humans to read and edit, and likewise
for machines.

The environment is a Python script which defines run-time variables, required by
Hadoop and Dumbo. Using a Python scripts for run-time variables leaves the
problem of validation to Python.

In order to join together all the different abilities which were wanted from the
data server, it was designed as a middleware application. Since the main goal
was not to build a middleware framework from scratch, an existing such was used.

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface} framework, it includes both middleware and application parts. The
middleware applications are responsible for dispatching requests to the
corresponding application or file.

@@

TODO: *this should go into implementation.*
The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.

% TODO: not sure this is precisely a story about the implementation.
% putting it here for now.

@@

\begin{verbatim}
  def map(key, value):
    log = parse_apache_log(value)
    dimensions = {'useragent' : log.useragent,
                  'path'      : log.path,
                  'status'    : log.statuscode}
    values = {'pageviews' : 1,
              'bytes'     : log.size}

    yield log.timestamp,dimensions,measurements
\end{verbatim}

@@

\section{Configuration}

\subsection{Data sets}

The user's data sets are defined using YAML. The data set files reside in
the project's \texttt{config} directory.

\subsection{Environment}

Certain paths need to be known by Zohmg in order to function, these are set
in the config/environment.py file.

In order to import data to Zohmg, the user writes a mapper which Zohmg wraps;
interpreting the output from the mapper program and storing it in HBase.

Support exists for writing the mapper in Java or Python.

@@

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed.

Python Paste is a framework for building WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
applications and middleware. \cite{definitive_guide_to_pylons}

@@

Middleware is software that acts as an intermediary. WSGI middleware
receives a WSGI request and then performs logic upon this request, before
passing the request to a WSGI application. \cite{paste}

@@

@@

@@

@@

% TODO: Why do we chose HBase in favour of other techniques (i.e. SQL)?

@@

% TODO: elaborate on reasons to why starsql does not scale

@@

>>>>

\documentclass[a4paper,10pt]{report}
\usepackage{graphicx}
\usepackage{listings}

@@

Summarize the method chapter.

Working methods.

Tools.

Something else-or-another.

@@

This chapter will present the implementation of the Zohmg system. For a usage
manual see the appendix.

% TODO: Do we need to talk about the storage implementation?

In the first section the implementation of project creation, configuration, and
setup, will be presented. The second section features the data import
implementation. The third section presents the implementation of data
interchange between Zohmg and other software but also throughout the Zohmg
system. The fourth, and final, section presents the implementation details of
the Zohmg system's exporting compontent.

\section{Project management}

A Zohmg project is necessary for using Zohmg; they are created with a management
script, which is also used for interacting with the Zohmg system. When the user
has created a Zohmg project directory further actions can be performed.  The

Zohmg projects reside within a project directory, which contains five
directories: \texttt{clients}, \texttt{config}, \texttt{lib}, \texttt{mappers},
and \texttt{transformers}. The details of these directories are presented
throughout this chapter.

\subsection*{Configuration}

The Zohmg system needs configuration of the user's data sets and the environment
for Hadoop and Dumbo. When the data set files are configured, the user runs the
management script, which executes the configuration and creates tables in HBase
according to the configuration.

\subsubsection*{Data sets}

In this context a data set is the schema that the user defines for wanted data
sets to be handled by Zohmg. The configuration of a data set is simply a list
of dimensions and projections, which will layout the HBase tables according to
the schema. This configuration will also be read during the import phase, for
handling data to be imported.

The data set files reside in the project's \texttt{config} directory. The data
sets are configured using YAML\footnote{YAML Ain't a Markup Language}, a
human-readable data serializaton format. Any human-readable configure file
method could have been chosen. YAML was choosen because it is is very simple
for humans to read and edit, and likewise for machines. The Python module
PyYAML was used to implement the YAML configuration reading part. \cite{pyyaml}

\subsubsection*{Environment}

The environment is a Python script which defines run-time variables, required by
Hadoop and Dumbo. Using a Python scripts for run-time variables leaves the
problem of validation to Python. The environment configuration file resides in
the project's \texttt{config} directory.

In order to import data to Zohmg, the user writes a map function in Python which
Zohmg wraps; interpreting the output from the mapper program and storing it in
HBase.

@@

\begin{lstlisting}[language=Python]
   def map(key, value):
      log = parse_apache_log(value)
      dimensions = {'useragent' : log.useragent,
                    'path'      : log.path,
                    'status'    : log.statuscode}
      values = {'pageviews' : 1,
                'bytes'     : log.size}

      yield log.timestamp,dimensions,measurements
\end{lstlisting}

@@

@@

In order to join together all the different abilities wanted from the data
server, it was designed as a middleware application. Middleware is software
that acts as an intermediary.

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
framework, which includes both middleware and application parts. The middleware
applications are responsible for dispatching requests to the corresponding
application or file.
WSGI middleware receives a
WSGI request and then performs logic upon this request, before passing the
request to a WSGI application. \cite{definitive_guide_to_pylons,paste}
Since the main goal was not
to build a middleware framework from scratch, an existing such was used.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed. \cite{python_paste}

@@

@@

Summary here.

@@

Discussion here.

@@

\section{Installation}

During the installation phase superuser (root) privileges are required.

Install Hadoop and HBase.

\begin{verbatim}
   sh scripts/install_hadoop_hbase.sh
\end{verbatim}

The above command will install Hadoop and HBase to \texttt{/opt} by default,
this is customizable with script arguments. Both Hadoop and HBase will be
configured for local (standalone) mode when the installation is done. Installing
Hadoop and HBase on a cluster requires additional attention, for
instance passphraseless SSH\footnote{Secure Shell, an encrypted data transport
protocol; it is commonly used for remote access.} keys and additional program
configuration.

When Hadoop and HBase is installed install Zohmg by invoking the following
command in the root of the source tree

\begin{verbatim}
   python install.py
\end{verbatim}

Creates eggs of Zohmg system and puts them in system-wide directory for python
packages.

Zohmg files copied, scripts, documentation, library files (other's software).

@@

TODO: Why do we chose HBase in favour of other techniques (i.e. SQL)?

@@

TODO: elaborate on reasons to why starsql does not scale

@@

\section{Electromagnetic field theory}

\section{Typical use case}

The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.
